### 2017-11-03 submission/manuscript.Rmd
add refs to background

Colorectal cancer (CRC) is a growing world wide health problem in which the microbiota has been purported to play an active role in disease pathogenesis. Numerous studies have shown the importance of both individual microbes () and the overall community () in polyp formation in mouse models. There has also been numerous case control studies investigating the microbiota in both adenoma and carcinoma. Recently, a meta-analysis was published investigating whether specific biomarkers could be consistently identified using multiple data sets (). Many of the studies along with the current meta-analysis focus on identifying biomarkers or individual microbes but do not critically investigate the community role in the disease.

Using both fecal and tissue samples totalling over 2100 total individuals across 14 studies within our data analysis we exapnd both the breadth and scope of the previous meta-analysis to investigate whether the bacterial community is an important risk factor for both adenoma and carcinoma. We first assessed the diversity of controls, adenoma, and carcinoma individuals and tested whether they change and if it results in an increased relative risk of adenoma or carcinoma. Next, we assessed how this relative risk compared to CRC associated genera for both adenoma and carcinoma. Third, using Random Forest models we assessed whether the community context can increase the classification model area under the curve (AUC). Finally, we examine whether the studies that were used were adequately powered and if not what effect size they were powered for.

>>>>

Colorectal cancer (CRC) is a growing world wide health problem [@siegel_cancer_2016] in which the microbiota has been purported to play an active role in disease pathogenesis [@flynn_metabolic_2016]. Numerous studies have shown the importance of both individual microbes [@goodwin_polyamine_2011; @abed_fap2_2016; @arthur_intestinal_2012; @kostic_fusobacterium_2013; @wu_human_2009] and the overall community [@zackular_manipulation_2016; @zackular_gut_2013; @baxter_structure_2014] in polyp formation in mouse models. There has also been numerous case control studies investigating the microbiota in both adenoma and carcinoma. Recently, a meta-analysis was published investigating whether specific biomarkers could be consistently identified using multiple data sets [@shah_leveraging_2017]. Many of the studies along with the current meta-analysis focus on identifying biomarkers or individual microbes but do not critically investigate the community role in the disease.

Using both fecal and tissue samples totalling over 2100 total individuals across 14 studies [@hale_shifts_2017; @flemer_tumour-associated_2017; @chen_human_2012; @wang_structural_2012; @lu_mucosal_2016; @brim_microbiome_2013; @weir_stool_2013; @dejea_microbiota_2014; @sanapareddy_increased_2012; @ahn_human_2013; @burns_virulence_2015; @zeller_potential_2014; @baxter_microbiota-based_2016; @geng_diversified_2013] within our data analysis we expand both the breadth and scope of the previous meta-analysis to investigate whether the bacterial community is an important risk factor for both adenoma and carcinoma. We first assessed the diversity of controls, adenoma, and carcinoma individuals and tested whether they change and if it results in an increased relative risk of adenoma or carcinoma. Next, we assessed how this relative risk compared to CRC associated genera for both adenoma and carcinoma. Third, using Random Forest models we assessed whether the community context can increase the classification model area under the curve (AUC). Finally, we examine whether the studies that were used were adequately powered and if not what effect size they were powered for.

### 2017-11-03 submission/manuscript.tex
add refs to background

Colorectal cancer (CRC) is a growing world wide health problem in which
the microbiota has been purported to play an active role in disease
pathogenesis. Numerous studies have shown the importance of both
individual microbes () and the overall community () in polyp formation
in mouse models. There has also been numerous case control studies
investigating the microbiota in both adenoma and carcinoma. Recently, a
meta-analysis was published investigating whether specific biomarkers
could be consistently identified using multiple data sets (). Many of
the studies along with the current meta-analysis focus on identifying
biomarkers or individual microbes but do not critically investigate the
community role in the disease.

Using both fecal and tissue samples totalling over 2100 total
individuals across 14 studies within our data analysis we exapnd both
the breadth and scope of the previous meta-analysis to investigate
whether the bacterial community is an important risk factor for both
adenoma and carcinoma. We first assessed the diversity of controls,
adenoma, and carcinoma individuals and tested whether they change and if
it results in an increased relative risk of adenoma or carcinoma. Next,
we assessed how this relative risk compared to CRC associated genera for
both adenoma and carcinoma. Third, using Random Forest models we
assessed whether the community context can increase the classification
model area under the curve (AUC). Finally, we examine whether the
studies that were used were adequately powered and if not what effect
size they were powered for.

@@

The driver-passenger model of the microbial role in CRC, as summarized
by Flynn {[}1{]}, can be supported with this data for carcinoma but not
necessarily for adenoma. The drasitically increased relative risk of
disease when considering the CRC associated genera is highly supportive
of this type of process. In a driver-passenger scenario it is possible
that simply having the driver present or only identifying the passenger
is a good enough proxy that the event is occuring. This would account
for the observation that there is no synergistic increase in relative
risk when accounting for either the total number or increasing abundance
of these genera. The initial establishment of the driver within the
system is also dependent on the community that is present and this is
supported by the observation that when adding the community context to
our models along with the CRC associated genera the model AUC increases.

@@

\textbf{\emph{Obtaining Data Sets:}} Studies used for this meta-analysis
were identified through the review articles written by Keku, et al. and
Vogtmann, et al. {[}2,3{]}. All studies were included that used tissue
or feces as their sample source for 16S rRNA gene sequencing analysis.
Studies using either 454 or Illumina sequencing technology were
included. Only data sets that had the raw sequences available for
analysis were included. Some studies did not have publically available
raw sequences or did not have meta data in which the authors were able
to share. After this filtering step the following studies remained: Ahn
{[}4{]}, Baxter {[}5{]}, Brim {[}6{]}, Burns {[}7{]}, Chen {[}8{]},
Dejea {[}9{]}, Flemer {[}10{]}, Geng {[}11{]}, Hale {[}12{]}, Kostic
{[}13{]}, Lu {[}14{]}, Sanapareddy {[}15{]}, Wang {[}16{]}, Weir
{[}17{]}, and Zeller {[}18{]}. The Zackular {[}19{]} study was not
included becasue the 90 individuals analyzed within the study are
contained within the larger Baxter study. The Kostic study was not used
since after sequence processing all the case samples did not have more
than 100 sequences remaining. This left a total of 13 studies in which
complete analysis could be completed.

@@

\textbf{\emph{Sequence Processing:}} For the majority of studies raw
sequences were downloaded from the SRA
(\url{ftp://ftp-trace.ncbi.nih.gov/sra/sra-instant/reads/ByStudy/sra/SRP/})
and metadata was obtained from the following website:
\url{http://www.ncbi.nlm.nih.gov/Traces/study/} by searching the
respective accession number of the study. Of the studies that did not
have sequences and meta data on the SRA one study had the data stored on
DBGap {[}4{]} and four studies the data was obtained directly from the
authors {[}10,12,15,17{]}. Each study was processed using the mothur
(v1.39.3) software program {[}20{]}. Where possible quality filtering
utilized the default methods used in mothur for either 454 or Illumina
based sequencing. If it was not possible to use these defaults the
author stated quality cut-offs were used instead. Chimeras were
identifed and removed using the VSEARCH {[}21{]} program and \emph{de
novo} OTU clustering at 97\% similarity using the OptiClust algorithm
{[}22{]} was utilized.

@@

\hypertarget{refs}{}
\hypertarget{ref-flynn_metabolic_2016}{}
1. Flynn KJ, Baxter NT, Schloss PD. Metabolic and Community Synergy of
Oral Bacteria in Colorectal Cancer. mSphere. 2016;1.

\hypertarget{ref-keku_gastrointestinal_2015}{}
2. Keku TO, Dulal S, Deveaux A, Jovov B, Han X. The gastrointestinal
microbiota and colorectal cancer. American Journal of Physiology -
Gastrointestinal and Liver Physiology {[}Internet{]}. 2015 {[}cited 2017
Oct 30{]};308:G351--63. Available from:
\url{http://ajpgi.physiology.org/lookup/doi/10.1152/ajpgi.00360.2012}

\hypertarget{ref-vogtmann_epidemiologic_2016}{}
3. Vogtmann E, Goedert JJ. Epidemiologic studies of the human microbiome
and cancer. British Journal of Cancer {[}Internet{]}. 2016 {[}cited 2017
Oct 30{]};114:237--42. Available from:
\url{http://www.nature.com/doifinder/10.1038/bjc.2015.465}

\hypertarget{ref-ahn_human_2013}{}
4. Ahn J, Sinha R, Pei Z, Dominianni C, Wu J, Shi J, et al. Human gut
microbiome and risk for colorectal cancer. Journal of the National
Cancer Institute. 2013;105:1907--11.

\hypertarget{ref-baxter_microbiota-based_2016}{}
5. Baxter NT, Ruffin MT, Rogers MAM, Schloss PD. Microbiota-based model
improves the sensitivity of fecal immunochemical test for detecting
colonic lesions. Genome Medicine. 2016;8:37.

\hypertarget{ref-brim_microbiome_2013}{}
6. Brim H, Yooseph S, Zoetendal EG, Lee E, Torralbo M, Laiyemo AO, et
al. Microbiome analysis of stool samples from African Americans with
colon polyps. PloS One. 2013;8:e81352.

\hypertarget{ref-burns_virulence_2015}{}
7. Burns MB, Lynch J, Starr TK, Knights D, Blekhman R. Virulence genes
are a signature of the microbiome in the colorectal tumor
microenvironment. Genome Medicine. 2015;7:55.

\hypertarget{ref-chen_human_2012}{}
8. Chen W, Liu F, Ling Z, Tong X, Xiang C. Human intestinal lumen and
mucosa-associated microbiota in patients with colorectal cancer. PloS
One. 2012;7:e39743.

\hypertarget{ref-dejea_microbiota_2014}{}
9. Dejea CM, Wick EC, Hechenbleikner EM, White JR, Mark Welch JL,
Rossetti BJ, et al. Microbiota organization is a distinct feature of
proximal colorectal cancers. Proceedings of the National Academy of
Sciences of the United States of America. 2014;111:18321--6.

\hypertarget{ref-flemer_tumour-associated_2017}{}
10. Flemer B, Lynch DB, Brown JMR, Jeffery IB, Ryan FJ, Claesson MJ, et
al. Tumour-associated and non-tumour-associated microbiota in colorectal
cancer. Gut. 2017;66:633--43.

\hypertarget{ref-geng_diversified_2013}{}
11. Geng J, Fan H, Tang X, Zhai H, Zhang Z. Diversified pattern of the
human colorectal cancer microbiome. Gut Pathogens. 2013;5:2.

@@

\hypertarget{ref-kostic_genomic_2012}{}
13. Kostic AD, Gevers D, Pedamallu CS, Michaud M, Duke F, Earl AM, et
al. Genomic analysis identifies association of Fusobacterium with
colorectal carcinoma. Genome Research. 2012;22:292--8.

\hypertarget{ref-lu_mucosal_2016}{}
14. Lu Y, Chen J, Zheng J, Hu G, Wang J, Huang C, et al. Mucosal
adherent bacterial dysbiosis in patients with colorectal adenomas.
Scientific Reports. 2016;6:26337.

\hypertarget{ref-sanapareddy_increased_2012}{}
15. Sanapareddy N, Legge RM, Jovov B, McCoy A, Burcal L, Araujo-Perez F,
et al. Increased rectal microbial richness is associated with the
presence of colorectal adenomas in humans. The ISME journal.
2012;6:1858--68.

\hypertarget{ref-wang_structural_2012}{}
16. Wang T, Cai G, Qiu Y, Fei N, Zhang M, Pang X, et al. Structural
segregation of gut microbiota between colorectal cancer patients and
healthy volunteers. The ISME journal. 2012;6:320--9.

\hypertarget{ref-weir_stool_2013}{}
17. Weir TL, Manter DK, Sheflin AM, Barnett BA, Heuberger AL, Ryan EP.
Stool microbiome and metabolome differences between colorectal cancer
patients and healthy adults. PloS One. 2013;8:e70803.

\hypertarget{ref-zeller_potential_2014}{}
18. Zeller G, Tap J, Voigt AY, Sunagawa S, Kultima JR, Costea PI, et al.
Potential of fecal microbiota for early-stage detection of colorectal
cancer. Molecular Systems Biology. 2014;10:766.

\hypertarget{ref-zackular_human_2014}{}
19. Zackular JP, Rogers MAM, Ruffin MT, Schloss PD. The human gut
microbiome as a screening tool for colorectal cancer. Cancer Prevention
Research (Philadelphia, Pa.). 2014;7:1112--21.

\hypertarget{ref-schloss_introducing_2009}{}
20. Schloss PD, Westcott SL, Ryabin T, Hall JR, Hartmann M, Hollister
EB, et al. Introducing mothur: Open-Source, Platform-Independent,
Community-Supported Software for Describing and Comparing Microbial
Communities. Appl.Environ.Microbiol. {[}Internet{]}. 2009 {[}cited 12AD
Jan 1{]};75:7537--41. Available from:
\url{http://aem.asm.org/cgi/content/abstract/75/23/7537}

\hypertarget{ref-rognes_vsearch_2016}{}
21. Rognes T, Flouri T, Nichols B, Quince C, Mahé F. VSEARCH: A
versatile open source tool for metagenomics. PeerJ. 2016;4:e2584.

\hypertarget{ref-westcott_opticlust_2017}{}
22. Westcott SL, Schloss PD. OptiClust, an Improved Method for Assigning
Amplicon-Based Sequence Data to Operational Taxonomic Units. mSphere.
2017;2.

>>>>

Colorectal cancer (CRC) is a growing world wide health problem {[}1{]}
in which the microbiota has been purported to play an active role in
disease pathogenesis {[}2{]}. Numerous studies have shown the importance
of both individual microbes {[}3--7{]} and the overall community
{[}8--10{]} in polyp formation in mouse models. There has also been
numerous case control studies investigating the microbiota in both
adenoma and carcinoma. Recently, a meta-analysis was published
investigating whether specific biomarkers could be consistently
identified using multiple data sets {[}11{]}. Many of the studies along
with the current meta-analysis focus on identifying biomarkers or
individual microbes but do not critically investigate the community role
in the disease.

Using both fecal and tissue samples totalling over 2100 total
individuals across 14 studies {[}12--25{]} within our data analysis we
expand both the breadth and scope of the previous meta-analysis to
investigate whether the bacterial community is an important risk factor
for both adenoma and carcinoma. We first assessed the diversity of
controls, adenoma, and carcinoma individuals and tested whether they
change and if it results in an increased relative risk of adenoma or
carcinoma. Next, we assessed how this relative risk compared to CRC
associated genera for both adenoma and carcinoma. Third, using Random
Forest models we assessed whether the community context can increase the
classification model area under the curve (AUC). Finally, we examine
whether the studies that were used were adequately powered and if not
what effect size they were powered for.

@@

The driver-passenger model of the microbial role in CRC, as summarized
by Flynn {[}2{]}, can be supported with this data for carcinoma but not
necessarily for adenoma. The drasitically increased relative risk of
disease when considering the CRC associated genera is highly supportive
of this type of process. In a driver-passenger scenario it is possible
that simply having the driver present or only identifying the passenger
is a good enough proxy that the event is occuring. This would account
for the observation that there is no synergistic increase in relative
risk when accounting for either the total number or increasing abundance
of these genera. The initial establishment of the driver within the
system is also dependent on the community that is present and this is
supported by the observation that when adding the community context to
our models along with the CRC associated genera the model AUC increases.

@@

\textbf{\emph{Obtaining Data Sets:}} Studies used for this meta-analysis
were identified through the review articles written by Keku, et al. and
Vogtmann, et al. {[}26,27{]}. All studies were included that used tissue
or feces as their sample source for 16S rRNA gene sequencing analysis.
Studies using either 454 or Illumina sequencing technology were
included. Only data sets that had the raw sequences available for
analysis were included. Some studies did not have publically available
raw sequences or did not have meta data in which the authors were able
to share. After this filtering step the following studies remained: Ahn
{[}21{]}, Baxter {[}24{]}, Brim {[}17{]}, Burns {[}22{]}, Chen {[}14{]},
Dejea {[}19{]}, Flemer {[}13{]}, Geng {[}25{]}, Hale {[}12{]}, Kostic
{[}28{]}, Lu {[}16{]}, Sanapareddy {[}20{]}, Wang {[}15{]}, Weir
{[}18{]}, and Zeller {[}23{]}. The Zackular {[}29{]} study was not
included becasue the 90 individuals analyzed within the study are
contained within the larger Baxter study. The Kostic study was not used
since after sequence processing all the case samples did not have more
than 100 sequences remaining. This left a total of 13 studies in which
complete analysis could be completed.

@@

\textbf{\emph{Sequence Processing:}} For the majority of studies raw
sequences were downloaded from the SRA
(\url{ftp://ftp-trace.ncbi.nih.gov/sra/sra-instant/reads/ByStudy/sra/SRP/})
and metadata was obtained from the following website:
\url{http://www.ncbi.nlm.nih.gov/Traces/study/} by searching the
respective accession number of the study. Of the studies that did not
have sequences and meta data on the SRA one study had the data stored on
DBGap {[}21{]} and four studies the data was obtained directly from the
authors {[}12,13,18,20{]}. Each study was processed using the mothur
(v1.39.3) software program {[}30{]}. Where possible quality filtering
utilized the default methods used in mothur for either 454 or Illumina
based sequencing. If it was not possible to use these defaults the
author stated quality cut-offs were used instead. Chimeras were
identifed and removed using the VSEARCH {[}31{]} program and \emph{de
novo} OTU clustering at 97\% similarity using the OptiClust algorithm
{[}32{]} was utilized.

@@

\hypertarget{refs}{}
\hypertarget{ref-siegel_cancer_2016}{}
1. Siegel RL, Miller KD, Jemal A. Cancer statistics, 2016. CA: a cancer
journal for clinicians. 2016;66:7--30.

\hypertarget{ref-flynn_metabolic_2016}{}
2. Flynn KJ, Baxter NT, Schloss PD. Metabolic and Community Synergy of
Oral Bacteria in Colorectal Cancer. mSphere. 2016;1.

\hypertarget{ref-goodwin_polyamine_2011}{}
3. Goodwin AC, Destefano Shields CE, Wu S, Huso DL, Wu X, Murray-Stewart
TR, et al. Polyamine catabolism contributes to enterotoxigenic
Bacteroides fragilis-induced colon tumorigenesis. Proceedings of the
National Academy of Sciences of the United States of America.
2011;108:15354--9.

\hypertarget{ref-abed_fap2_2016}{}
4. Abed J, Emgård JEM, Zamir G, Faroja M, Almogy G, Grenov A, et al.
Fap2 Mediates Fusobacterium nucleatum Colorectal Adenocarcinoma
Enrichment by Binding to Tumor-Expressed Gal-GalNAc. Cell Host \&
Microbe. 2016;20:215--25.

\hypertarget{ref-arthur_intestinal_2012}{}
5. Arthur JC, Perez-Chanona E, Mühlbauer M, Tomkovich S, Uronis JM, Fan
T-J, et al. Intestinal inflammation targets cancer-inducing activity of
the microbiota. Science (New York, N.Y.). 2012;338:120--3.

\hypertarget{ref-kostic_fusobacterium_2013}{}
6. Kostic AD, Chun E, Robertson L, Glickman JN, Gallini CA, Michaud M,
et al. Fusobacterium nucleatum potentiates intestinal tumorigenesis and
modulates the tumor-immune microenvironment. Cell Host \& Microbe.
2013;14:207--15.

\hypertarget{ref-wu_human_2009}{}
7. Wu S, Rhee K-J, Albesiano E, Rabizadeh S, Wu X, Yen H-R, et al. A
human colonic commensal promotes colon tumorigenesis via activation of T
helper type 17 T cell responses. Nature Medicine. 2009;15:1016--22.

\hypertarget{ref-zackular_manipulation_2016}{}
8. Zackular JP, Baxter NT, Chen GY, Schloss PD. Manipulation of the Gut
Microbiota Reveals Role in Colon Tumorigenesis. mSphere. 2016;1.

\hypertarget{ref-zackular_gut_2013}{}
9. Zackular JP, Baxter NT, Iverson KD, Sadler WD, Petrosino JF, Chen GY,
et al. The gut microbiome modulates colon tumorigenesis. mBio.
2013;4:e00692--00613.

\hypertarget{ref-baxter_structure_2014}{}
10. Baxter NT, Zackular JP, Chen GY, Schloss PD. Structure of the gut
microbiome following colonization with human feces determines colonic
tumor burden. Microbiome. 2014;2:20.

\hypertarget{ref-shah_leveraging_2017}{}
11. Shah MS, DeSantis TZ, Weinmaier T, McMurdie PJ, Cope JL, Altrichter
A, et al. Leveraging sequence-based faecal microbial community survey
data to identify a composite biomarker for colorectal cancer. Gut. 2017;

@@

\hypertarget{ref-flemer_tumour-associated_2017}{}
13. Flemer B, Lynch DB, Brown JMR, Jeffery IB, Ryan FJ, Claesson MJ, et
al. Tumour-associated and non-tumour-associated microbiota in colorectal
cancer. Gut. 2017;66:633--43.

\hypertarget{ref-chen_human_2012}{}
14. Chen W, Liu F, Ling Z, Tong X, Xiang C. Human intestinal lumen and
mucosa-associated microbiota in patients with colorectal cancer. PloS
One. 2012;7:e39743.

\hypertarget{ref-wang_structural_2012}{}
15. Wang T, Cai G, Qiu Y, Fei N, Zhang M, Pang X, et al. Structural
segregation of gut microbiota between colorectal cancer patients and
healthy volunteers. The ISME journal. 2012;6:320--9.

\hypertarget{ref-lu_mucosal_2016}{}
16. Lu Y, Chen J, Zheng J, Hu G, Wang J, Huang C, et al. Mucosal
adherent bacterial dysbiosis in patients with colorectal adenomas.
Scientific Reports. 2016;6:26337.

\hypertarget{ref-brim_microbiome_2013}{}
17. Brim H, Yooseph S, Zoetendal EG, Lee E, Torralbo M, Laiyemo AO, et
al. Microbiome analysis of stool samples from African Americans with
colon polyps. PloS One. 2013;8:e81352.

\hypertarget{ref-weir_stool_2013}{}
18. Weir TL, Manter DK, Sheflin AM, Barnett BA, Heuberger AL, Ryan EP.
Stool microbiome and metabolome differences between colorectal cancer
patients and healthy adults. PloS One. 2013;8:e70803.

\hypertarget{ref-dejea_microbiota_2014}{}
19. Dejea CM, Wick EC, Hechenbleikner EM, White JR, Mark Welch JL,
Rossetti BJ, et al. Microbiota organization is a distinct feature of
proximal colorectal cancers. Proceedings of the National Academy of
Sciences of the United States of America. 2014;111:18321--6.

\hypertarget{ref-sanapareddy_increased_2012}{}
20. Sanapareddy N, Legge RM, Jovov B, McCoy A, Burcal L, Araujo-Perez F,
et al. Increased rectal microbial richness is associated with the
presence of colorectal adenomas in humans. The ISME journal.
2012;6:1858--68.

\hypertarget{ref-ahn_human_2013}{}
21. Ahn J, Sinha R, Pei Z, Dominianni C, Wu J, Shi J, et al. Human gut
microbiome and risk for colorectal cancer. Journal of the National
Cancer Institute. 2013;105:1907--11.

\hypertarget{ref-burns_virulence_2015}{}
22. Burns MB, Lynch J, Starr TK, Knights D, Blekhman R. Virulence genes
are a signature of the microbiome in the colorectal tumor
microenvironment. Genome Medicine. 2015;7:55.

\hypertarget{ref-zeller_potential_2014}{}
23. Zeller G, Tap J, Voigt AY, Sunagawa S, Kultima JR, Costea PI, et al.
Potential of fecal microbiota for early-stage detection of colorectal
cancer. Molecular Systems Biology. 2014;10:766.

\hypertarget{ref-baxter_microbiota-based_2016}{}
24. Baxter NT, Ruffin MT, Rogers MAM, Schloss PD. Microbiota-based model
improves the sensitivity of fecal immunochemical test for detecting
colonic lesions. Genome Medicine. 2016;8:37.

\hypertarget{ref-geng_diversified_2013}{}
25. Geng J, Fan H, Tang X, Zhai H, Zhang Z. Diversified pattern of the
human colorectal cancer microbiome. Gut Pathogens. 2013;5:2.

\hypertarget{ref-keku_gastrointestinal_2015}{}
26. Keku TO, Dulal S, Deveaux A, Jovov B, Han X. The gastrointestinal
microbiota and colorectal cancer. American Journal of Physiology -
Gastrointestinal and Liver Physiology {[}Internet{]}. 2015 {[}cited 2017
Oct 30{]};308:G351--63. Available from:
\url{http://ajpgi.physiology.org/lookup/doi/10.1152/ajpgi.00360.2012}

\hypertarget{ref-vogtmann_epidemiologic_2016}{}
27. Vogtmann E, Goedert JJ. Epidemiologic studies of the human
microbiome and cancer. British Journal of Cancer {[}Internet{]}. 2016
{[}cited 2017 Oct 30{]};114:237--42. Available from:
\url{http://www.nature.com/doifinder/10.1038/bjc.2015.465}

\hypertarget{ref-kostic_genomic_2012}{}
28. Kostic AD, Gevers D, Pedamallu CS, Michaud M, Duke F, Earl AM, et
al. Genomic analysis identifies association of Fusobacterium with
colorectal carcinoma. Genome Research. 2012;22:292--8.

\hypertarget{ref-zackular_human_2014}{}
29. Zackular JP, Rogers MAM, Ruffin MT, Schloss PD. The human gut
microbiome as a screening tool for colorectal cancer. Cancer Prevention
Research (Philadelphia, Pa.). 2014;7:1112--21.

\hypertarget{ref-schloss_introducing_2009}{}
30. Schloss PD, Westcott SL, Ryabin T, Hall JR, Hartmann M, Hollister
EB, et al. Introducing mothur: Open-Source, Platform-Independent,
Community-Supported Software for Describing and Comparing Microbial
Communities. Appl.Environ.Microbiol. {[}Internet{]}. 2009 {[}cited 12AD
Jan 1{]};75:7537--41. Available from:
\url{http://aem.asm.org/cgi/content/abstract/75/23/7537}

\hypertarget{ref-rognes_vsearch_2016}{}
31. Rognes T, Flouri T, Nichols B, Quince C, Mahé F. VSEARCH: A
versatile open source tool for metagenomics. PeerJ. 2016;4:e2584.

\hypertarget{ref-westcott_opticlust_2017}{}
32. Westcott SL, Schloss PD. OptiClust, an Improved Method for Assigning
Amplicon-Based Sequence Data to Operational Taxonomic Units. mSphere.
2017;2.
